MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-25 chapters/MetaTracts.tex
Local text changes

In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength, high corrosion resistance. CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
At the same time highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates, consist of two main components:
A matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles. 
The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component.

Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. Typical scanning geometry in industrial XCT is the specimen placed on a rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays get attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

@@

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx} 
A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Basser2000, Mori1999, Mori2002}. Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion. 
 This is usually done by solving a second-order Runge-Kutta integration. The principal diffusion direction at each discrete location is interpolated to from a continuous velocity field. 
 Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2002},~\cite{Basser2000} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data. 
Mori et al.~\cite{Mori1999},~\cite{Mori2002} showed that tract reconstruction techniques could be broadly divided into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.

@@

Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}) can be applied to the problem of organizing fiber trace data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber traces. In the ideal case fiber traces that belong to the same bundle must remain ``close" to each other in the lower dimensional space. Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of pointwise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}.

The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationship. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results.

There are a number of difficulties in hierarchical clustering. First, computing all pairs distances for tracts to generate the distance matrix is time consuming (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times). Second, a ``correct'' distance measure to compare tracts must be chosen. Thirdly, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

@@

\subsection {Visual Analysis of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores: Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. This approach allows analyzing fiber orientations based on direction transfer functions. Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations. Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites. Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Like~\cite{Weissenbock2014}, most carbon fiber extraction literature deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.

Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes has received extensive research and recently the application of similar techniques to fiber tracing in composite materials has gained momentum. To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\caption{Data characteristics: (A) shows a rendering of the data. (B) shows 2D slices along the X- and (C) along the Z-axis. (D) close up of the region marked in green in (C), multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$, the data type is unsigned 8 bit integer. Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}a shows a volume rendering of the dataset. Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis. Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}d contains two bundles going in opposite directions. However, the separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern the fiber bundles cross each other in different orientations. In general the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.

@@

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}[tb]
\centering
	\includegraphics[width=0.5\textwidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}A, clustered into 10 clusters. The unit for length is the grid cube length. Fig.~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6. Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests. A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. On the other hand a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests. $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays, we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited, thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well. We compared  $m=3-7$ experimentally without any dramatic change in results. 
\begin{figure}[tb]
	\centering
	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
	\label{fig:comparison}
\end{figure}

>>>>

In general, CFRP materials have desirable characteristics such as high specific stiffness, high specific strength and high corrosion resistance. CFRP materials have these characteristics  at considerably lower weight compared to conventional materials.
At the same time, highly complex and integrated components which were previously impossible to manufacture may be produced from CFRPs.
Primary structures and highly loaded components in aeronautics are one example.

Typically, carbon fiber reinforced polymer components, more specifically CFRP laminates, consist of two main components:
a matrix on one side, which acts as a bonding component, and  reinforcements on the other side, which allow for achieving the desired characteristics.  Various production processes are used to manufacture CFRP laminates with endless carbon fibers. Most of these processes start with the reinforcement component by weaving individual carbon fiber bundles (yarn) into sheets of carbon fiber cloth in a predefined pattern. These sheets of woven carbon fiber cloth, which form the geometrical structure of the final CFRP materials, are also referred to as fabric. Depending on the requirements of the final component as well as on the manufacturing process, fabrics may be stacked in multiple layers of the cloth. Both alignment of fabrics and the weaving pattern of the individual carbon fiber bundles strongly influence the strength, durability and stiffness properties. Resins are integrated in the material system to fill the gaps in the fabric, forming the matrix component. The main function of this matrix component is to act as a bonding between the individual carbon fiber bundles. 
The increasing share of CFRPs as well as the complexity of material system and final components generated a strong demand for non-destructive testing (NDT) techniques for quality control~\cite{Red2012}. Ultrasonic testing is the most commonly used method for non-destructive testing of carbon fiber reinforce polymers. On one hand, ultrasonic testing provides a quick and cost-efficient overview of the material. On the other hand, it shows low resolution and may generate arbitrary results, e.g. due to the geometry of the final component.

Industrial 3D X-ray computed tomography (XCT, also referred to as 3DXCT or cone beam XCT) is increasingly applied for non-destructive testing of fiber reinforced polymers~\cite{Kastner2012}. XCT generates a 3D volumetric representation of the scanned specimen, reconstructed from a series of 2D penetration images, which are taken throughout a full rotation of the specimen. Typical scanning geometry in industrial XCT is the specimen placed on a rotary table between X-ray source and detector. X-rays are emitted by the source in the form of a conic beam. When passing through the specimen the X-rays are attenuated by the material in the specimen. The detector records the corresponding 2D attenuation image (penetration image) by transferring the X-rays in a scintillator layer into visible light. To generate the required series of 2D penetration images of a full $360^\circ$ rotation for a complete reconstruction of the data volume, the specimen is rotated stepwise and at each position a cone beam penetration image is recorded~\cite{heinzl-2008-thesis}. Finally the series of penetration images is  reconstructed as a 3D volumetric dataset of the specimen. 
In state of the art devices, XCT can reach voxel sizes below 500 $nm$. XCT generates high resolution volume data for comprehensive and detailed analyses of the test specimens. Unfortunately, there is still a trade-off between viewport and image resolution in XCT using cone beam geometry. The magnification reached within an XCT scan is determined by the specified distances between source and specimen as well as source and detector. The magnification therefore directly influences both resolution and viewport: higher resolutions decrease the viewport but show more details while lower resolutions allow for larger viewports and thus larger portions of the specimen.

@@

\subsection {Fiber Tracking in Diffusion Tensor Imaging}
\label{subsec:fiberEx} 
A basic assumption in DTI analysis is that the principal eigenvector of the diffusion tensor is parallel to the underlying dominant fiber direction in each image voxel~\cite{Basser2002, Basser2000, Mori1999, Mori2002}. Continuous tracts are created by propagating virtual particles along fiber directions until they reach some termination criterion. 
 This is usually done by solving a second-order Runge-Kutta integration. The principal diffusion direction at each discrete location is interpolated to form a continuous velocity field. 
 Because decisions are made locally, these methods perform poorly in noisy regions and often generate small fibers. Basser et al.~\cite{Basser2002,Basser2000} proposed that white matter tracts could be represented as 3D curves in space. They showed that numerical methods could be used to follow fibers and fiber bundles and to generate tracts in human brain data. 
Mori et al.~\cite{Mori1999,Mori2002} divided reconstruction techniques into line-propagation or energy minimization techniques. In line propagation approaches, trajectories are computed based on local neighborhoods and in energy minimization approaches the most  favorable trajectory connecting two given endpoints is selected.

@@

Brun et al.~\cite{Brun2003} described how a spectral non-linear dimensionality reduction technique,  Laplacian eigenmaps (Belkin and Niyogi~\cite{Belkin01}), can be applied to the problem of organizing fiber tracts data. The key notion of the Laplacian eigenmaps algorithm is to represent the underlying data as a graph. Each node represents a data point and the edges connect neighboring data points. An eigenvalue problem is solved to represent the data in a lower dimensional space while preserving the local graph structure. In the case of fiber bundles, the individual points are fiber tracts. In the ideal case fiber tracts that belong to the same bundle must remain ``close" to each other in the lower dimensional space. Westin et al.~\cite{westinMEDIA02} also use spectral clustering on a Hausdorff distance measure defined as the maximum of point wise minimum distances between two fibers. Jonasson et al.~\cite{jonasson2005} run k-means clustering on the eigenvectors of the affinity matrix defined as the co-occurrence of fibers in the data.
%The clustering methods have also been evaluated to be generally accurate in capturing fiber bundles as seen in Moberts et al.~\cite{Moberts2005} and Corouge et al.~\cite{Corouge2004}.

The agglomerative hierarchical clustering method \cite{DudaHartStork01} has gained popularity for proximity based fiber segregation (Zhang et al.~\cite{Zhang2008}, Corouge et al.~\cite{Corouge2004}). 
These approaches build on the assumption that proximity measures that compare DTI fiber trajectories can also represent anatomical relationships. An agglomerative hierarchical clustering method starts with each data point/fiber in an individual cluster. At each stage of the algorithm the two most similar clusters are merged based on some criterion. The two basic cluster similarity measures are single-link and complete-link. With the single link measure the distance between the clusters is the distance between the closest pair of items. 
Moberts et al.~\cite{Moberts2005} implemented several distance measures in their evaluation of fiber clustering and concluded that clustering methods are generally accurate in capturing fiber bundles. 
%Mean of closest distances performs better than closest point distance computations or Hausdorff distance measures and hierarchical clustering using single-link measure gave the best results.

There are a number of difficulties in hierarchical clustering. First, computing all pairs distances for tracts to generate the distance matrix is time consuming (see~\cite{Garyfallidis2012} for a recent work on decreasing computation times). Second, a ``correct'' distance measure to compare tracts must be chosen. Third, hierarchical clustering is best suited for similar length fibers.
Spectral methods are also hindered by long matrix computations.

@@

\subsection {Visual Analysis of Fiber Reinforced Composites}
The approaches presented in visualization and analysis of composites mainly focus on individual objects such as fibers or pores. Fritz et al.~\cite{Fritz2009} proposed interactive workflows for non destructive testing practitioners to explore and quantify steel fibers in reinforced sprayed concrete. This approach allows analyzing fiber orientations based on direction transfer functions. Salaberger et al.~\cite{Salaberger2011} introduced a pipeline to extract and characterize individual fibers of fiber reinforced composites. They encode the extracted fibers as color-coded line segments in 3D and visually identifying fibers with similar orientations. Reh et al.~\cite{PMI_AR_2012} introduced an approach to explore pores of carbon fiber reinforced composites. Recently, Weissenbock et al.~\cite{Weissenbock2014} introduced a system for interactive exploration and analysis of fibers in fiber reinforced polymers. 
Like~\cite{Weissenbock2014}, most carbon fiber extraction literature deals with fiber extraction from high resolution data where the individual fibers are clearly discernible.

Tools based on image analysis, fiber tracking and clustering into bundles in DTI and their subsequent visualization through stream tubes and coloring schemes have received extensive research. Recently, similar techniques have been applied to fiber tracing in composite materials. To the best of our knowledge there is no approach focusing on direct extraction of fiber bundle structures from low resolution XCT of composite fibers, which is the main scope of this work.

@@

\begin{figure}[tb]
\centering
%\includegraphics[width=0.45\textwidth]{imagesMT2014/MT_data_4_2.png}
\includegraphics[width=0.5\textwidth, trim = 0mm 110mm 0mm 0mm, clip,]{images_pvis/figure1}
\caption{Data characteristics: (a) Rendering of the data. (b) 2D slices along the X-axis and (c) 2D slice along the Z-axis. (d) Magnified region marked in green square (c). Multiple fiber bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}
Figure \ref{fig:data-char} shows dataset 1 with woven fiber bundles. Dataset 1 is 450$\times$300$\times$500 voxels in size with isotropic resolution of 2$\mu m$. The data type is unsigned 8 bit integer. Figure~\ref{fig:data-char} clearly shows the weaving structure of the composite unit cell as regional subset of the recurring fiber bundle weaving pattern. The unit cell is used as a basis structure for CFRP manufacturing. The alignment and weaving pattern of the fiber bundles in the cloth influences the strength and stiffness properties of the resulting material. Figure~\ref{fig:data-char}a shows a volume rendering of the dataset. Figure~\ref{fig:data-char}b shows 2D slices along the $X$- and Figure~\ref{fig:data-char}c in $Z$-axis. Figure~\ref{fig:data-char}d shows the magnified version of the green region of interest. 
The resolution of dataset 1 renders the individual carbon fibers in a fiber bundle hard to resolve using 2D slice images only. Figure~\ref{fig:data-char}d contains two bundles going in opposite directions. The separation between two fiber bundles is barely visible. The carbon fibers themselves do not differ much from the underlying epoxy matrix in terms of attenuation which poses an additional challenge for characterization. The fiber bundles may differ in terms of the amount of fibers in the bundle. Figure~\ref{fig:data-char}a shows the large variation in cross section sizes among the bundles.
Depending on the weaving pattern, the fiber bundles cross each other in different orientations. In general, the number of orientations is limited by the weaving pattern. Nevertheless, the weaving pattern may cause the individual bundles to be curved. In consequence individual fibers may be adjacent in Euclidean distance but belong to different bundles.

@@

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}[tb]
\centering
	\includegraphics[width=0.5\textwidth,  trim = 0mm 260mm 0mm 15mm, clip]{images_pvis/figure9}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}A, clustered into 10 clusters. The unit for length is the grid cube length. Fig.~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6. Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{h}$ to be 0.3 for all our tests. A $R_{h}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. On the other hand a very high $ R_{h}$ would lead to be a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{h}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the width parameters for the cylinders for the MetaTracts decide how coarse the fiber cylinders are. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and breadth(measured in grid cube length), respectively for all tests. $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays, we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited, thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well. We compared  $m=3-7$ experimentally without any dramatic change in results. 
%\begin{figure}[tb]
%	\centering
%	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
%	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
%	\label{fig:comparison}
%\end{figure}
%

### 2014-09-25 metaTracts-pvis2014.tex
local commit
reviewed user study, modified enums + itemizes
Local text changes

\documentclass{vgtc}                          % final (conference style)
%\documentclass[review]{vgtc}                 % review
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint]{vgtc}               % preprint
%\documentclass[electronic]{vgtc}             % electronic version

@@

\let\ifpdf\relax
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{amssymb}
\usepackage{subfig}
\usepackage{wrapfig}
\usepackage{comment}
\usepackage{epstopdf}
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

@@

%% Abstract section.
\abstract{
This work introduces MetaTracts, a novel method  for extracting and visualizing individual fiber bundles as well as their weaving patterns from X-ray computed tomography (XCT) scans of  endless carbon fiber reinforced  polymers (CFRP). The  proposed work flow is designed  to analyze large areas of CFRP material  such as complete unit cells integrating the recurring weaving pattern. It is designed to handle XCT scans of low resolution, in which  individual  fibers  are not  visible or are barely visible. 
First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
In the second step these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach: first by orientation then by proximity.
The tool can also  generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
We evaluate the proposed work flow on a number of real world datasets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
%mention user study here?%
%This work will provide a stepping stone to introduce novel non destructive testing techniques for reinforced fiber bundles analysis.
} % end of abstract

>>>>

%\documentclass{vgtc}                          % final (conference style)
\documentclass[review]{vgtc}                 % review
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint]{vgtc}               % preprint
%\documentclass[electronic]{vgtc}             % electronic version

@@

\let\ifpdf\relax
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}
\usepackage{amssymb}
\usepackage{subfig}
\usepackage{wrapfig}
\usepackage{comment}
\usepackage{epstopdf}
\usepackage{enumitem}
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

@@

%% Abstract section.
\abstract{
This work introduces MetaTracts, a novel method  for extracting and visualizing individual fiber bundles and weaving patterns from X-ray computed tomography (XCT) scans of  endless carbon fiber reinforced  polymers (CFRP). The  proposed work flow is designed  to analyze large areas of CFRP material  such as complete unit cells integrating the recurring weaving pattern. It is designed to handle XCT scans of low resolution, in which  individual  fibers  are not  visible or are barely visible. 
First, a coarse version of integral curves is used to trace subsections of the individual fiber bundles in the woven CFRP materials. We call these sections MetaTracts.
In the second step, these extracted fiber bundle sections (MetaTracts) are clustered using a two step approach: first by orientation, then by proximity.
The tool can generate volumetric representations as well as surface models of the extracted fiber bundles to be exported for further analysis.
We evaluate the proposed work flow on a number of real world datasets and demonstrate that MetaTracts effectively and robustly identifies and separates different fiber bundles. 
%mention user study here?%
%This work will provide a stepping stone to introduce novel non destructive testing techniques for reinforced fiber bundles analysis.
} % end of abstract
